---
title: "Class 5: Data Visuaization with GGPLOT"
author: "Alma"
format: html
toc: true
---

# Our first plot 

R has base graphics 


```{r}
head (cars)
```

```{r}
plot (cars)
```

How would I plot this with `ggplot2`? 
No, need to install using `install.packages()` function. Don't install in the code chunk since you will render the function every time. You can add a `#` 

Before I can use this package I need to load with a `library()` call.

```{r}
library(ggplot2)
ggplot(cars)
```
Every ggplot needs at least 3 layers:

-**Data** (i.e. the data frame we have)

-**Aesthetic** (the aesthetic mapping of our data to what we want to plot)

-**Geoms**(how we want to plot this stuff)

```{r}
ggplot(data=cars) +
  aes(x=speed,y=dist) +
geom_smooth(method = lm, se = FALSE) + geom_point()
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
```

```{r}
genes <- read.delim(url)
```

```{r}
head(genes)
```
Q1. how many genes are in this dataset?

```{r}
nrow(genes)
```

There are `r nrow(genes)` genes in this data set. 

```{r}
colnames(genes)
```

```{r}
table(genes$State)
```

```{r}
p <- ggplot(genes) +
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point()
```

```{r}
p
```

```{r}
c <- p + scale_colour_manual( values=c("blue","gray","red") ) 

c
```

Alpha has to be in the geom_point to make the points transparent 

To change the axis labels from Condition1 to Control, you have to use `lab(x = "Control")`

```{r}
c + labs(x = "Control (No Drug)", y = "Drug Treatment")
```

```{r}
library(plotly)

ggplotly(c)
```


```{r}
# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```

```{r}
gapminder 
head(gapminderin)
```

```{r}
# install.packages ('dplyr')

library(dplyr)
gapminder_2007 <- gapminder %>% filter(year==2007)
```

```{r}
ggplot(gapminder_2007) +
    aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha=0.5)
```


