---
title: "Class 10: Halloween Mini-Project"
author: "Alma Renero"
format: pdf
---

#Background 

In this mini-project we will examine 538 halloween candy data. What is your favorite candy? what is nougat anyway? and how do you say it in America?

First step is to read the data...

```{r}
candy <- read.csv("candy-data.txt", row.names=1)
head(candy)
```

How many candies there is?
```{r}
nrow(candy)
```
how many fruity candy types are in th data base?
```{r}
sum(candy$fruity)
```

```{r}
rownames(candy)
```

Q3. What is your favorite candy in the dataset and what is it’s winpercent value?
```{r}
candy["Sour Patch Kids",]$winpercent
```

Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",]$winpercent
```

Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?
```{r}
candy["Toosie Roll Snack Bars",]$winpercent
```

```{r}
skimr::skim(candy)
```
Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
```{r}

```


Q7. What do you think a zero and one represent for the candy$chocolate column?
```{r}
candy$chocolate
```

```{r}
hist(candy$winpercent,breaks=30)
```

```{r}
library("ggplot2")

ggplot(candy) + 
  aes(winpercent)+
  geom_histogram(bins=8, col="red", fill="yellow")
```

Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate.inds <- as.logical(candy$chocolate)
chocolate.win <- candy[chocolate.inds,]$winpercent
mean(chocolate.win)

```

```{r}
fruity.inds <- as.logical(candy$fruity)
fruity.win <- candy[fruity.inds,]$winpercent
mean(fruity.win)
```

Q12. Is this difference statistically significant?
```{r}
t.test(chocolate.win,fruity.win)
```

#3 Overall Candy Rankings

The base R `sort()` and `order()` functions are very useful!

```{r}
x <- c(5,1,2,6)
sort(x)
```

```{r}
x[order(x)]
```

```{r}
y <- c("barry","alice","chandra")
y
sort(y)
order(y)
```

First I want to order/arrange the whole database by winpercent values
```{r}
inds <- order(candy$winpercent)
head(candy[inds,],n=5)
```

Q15 make a barplot 

The default barplot, made with `geom_col()` has the bars 
```{r}
ggplot(candy)+
  aes(winpercent,row.names(candy),winpercent) +
  geom_col()
```

```{r}
ggplot(candy)+
  aes(winpercent, reorder( rownames(candy),winpercent)) +
  geom_col()
```

```{r}
ggsave("mybarplot.png")
```

Lets setup a color vector (that signifies candy type) that we can then use for some
```{r}
my_cols <- rep("black", nrow(candy))
#my_cols
my_cols[as.logical(candy$chocolate)] <- "chocolate"
my_cols[as.logical(candy$bar)] <- "brown"
my_cols[as.logical(candy$fruity)] <- "blue"
my_cols
```

Use this vector 

```{r}
ggplot(candy)+
  aes(winpercent, reorder( rownames(candy),winpercent)) +
  geom_col(fill=my_cols)
```
#5 Exploring the correlation structure 

What about the value for money?
what is the best candy for the least money?

One way to get at this would be to make 
```{r}
ggplot(candy)+
  aes(winpercent, pricepercent, label=rownames(candy))+
  geom_point(col=my_cols) +
  geom_text(col=my_cols)
```
```{r}
library(ggrepel)
```

```{r}
ggplot(candy)+
  aes(winpercent, pricepercent, label=rownames(candy))+
  geom_point(col=my_cols) +
  geom_text_repel(col=my_cols, max.overlaps = 7)
```
```{r}
library(corrplot)

cij <- cor(candy)
corrplot(cij)
```

#PCA Principal Component Analysis

The main function that always there for us is `prcomp()`. It has an important argument that is set to `scale=false`

```{r}
pca <- prcomp(candy, scale=TRUE)
summary (pca)
```

PCA plot = PC1 V. PC2
```{r}
plot(pca$x[-21], col=my_cols, pch=16)

```
Nicer plot . Ggplot only works with data,frames, as input so I need 
my_data=
```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```


```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```


```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (blue), other (black)",
       caption="Data from 538")
```

